Concurrent Programming with CXL

### 2020-06-19 book/paper.tex
Latest

CXL is not an object-oriented language like Python is.  In Python
you can pass a reference to an object to a method, and that method
can then update the object.  In CXL, it is also sometimes convenient
to have a method update a shared variable specified as an argument.
For this (as well as some other uses), CXL supports \emph{pointers}
\index{pointer}
to shared variables.
If \texttt{x} is a shared variable, then the expression \texttt{\&(x)}
(the parentheses are mandatory) is a pointer to \texttt{x}
(also known as the \emph{address}
\index{address}
of \texttt{x}).
Conversely, if \texttt{p} is a pointer to a shared variable, then the
expression \texttt{\^{}p} is the value of the shared variable.

Figure~\ref{fig:petersonmethods} again shows Peterson's algorithm,
but this time with methods defined to enter and exit the critical
section.
You can put the first three methods in its own CXL source file
and include it using the CXL \texttt{import} statement.
\index{import statement}
\index{module}
This would
make the code re-usable by other applications.  For those who are
extra adventurous: you can add the \texttt{P\_enter} and
\texttt{P\_exit} methods to the \texttt{P\_mutex} dictionary
like so:
\begin{code}
\begin{verbatim}
dict{ .turn: 0, .flags: [ False, False ], .enter: P_enter, .exit: P_exit }
\end{verbatim}
\end{code}
That would allow you to simulate object methods.

@@

While the CVM does not have any built-in interlock instructions,
it does have support for executing multiple instructions atomically.
This feature is available in the CXL language in two ways.
First, any CXL statement can be made atomic by placing a label in front
of it.  Second, a group of CXL statements can be made atomic
through its \texttt{atomic}
\index{atomic statement}
statement.
We can use \texttt{atomic} blocks to implement a wide variety of
interlock operations.
For example, we could fix the program in Figure~\ref{fig:inc} by
constructing an atomic increment operation for a counter, like so:
\begin{code}
\begin{Verbatim}[xleftmargin=5mm,numbers=left]
def atomic_inc(ptr):
    atomic:
        ^ptr = ^ptr + 1;
    ;
;
count = 0;
atomic_inc &(count);
\end{Verbatim}
\end{code}

>>>>

CXL is not an object-oriented language like Python is.  In Python
you can pass a reference to an object to a method, and that method
can then update the object.  In CXL, it is also sometimes convenient
to have a method update a shared variable specified as an argument.
For this (as well as some other uses), CXL supports \emph{pointers}
\index{pointer}
to shared variables.
If \texttt{x} is a shared variable, then the expression \texttt{\&x}
 is a pointer to \texttt{x} (also known as the \emph{address}
\index{address}
of \texttt{x}).
Conversely, if \texttt{p} is a pointer to a shared variable, then the
expression \texttt{\^{}p} is the value of the shared variable.
In particular, \texttt{\^{}\&x == x}.
This is similar to how C pointers work.

Figure~\ref{fig:petersonmethods} again shows Peterson's algorithm,
but this time with methods defined to enter and exit the critical
section.
You can put the first three methods in its own CXL source file
and include it using the CXL \texttt{import} statement.
\index{import statement}
\index{module}
This would
make the code re-usable by other applications.

For those who are
extra adventurous: you can add the \texttt{P\_enter} and
\texttt{P\_exit} methods to the \texttt{P\_mutex} dictionary
like so:
\begin{code}
\begin{verbatim}
dict{ .turn: 0, .flags: [ False, False ], .enter: P_enter, .exit: P_exit }
\end{verbatim}
\end{code}
That would allow you to simulate object methods.

@@

While the CVM does not have any built-in interlock instructions,
it does have support for executing multiple instructions atomically.
This feature is available in the CXL language in two ways.
First, any CXL statement can be made atomic by placing a label in front
of it.  Second, a group of CXL statements can be made atomic
through its \texttt{atomic}
\index{atomic statement}
statement.
We can use \texttt{atomic} blocks to implement a wide variety of
interlock operations.
For example, we could fix the program in Figure~\ref{fig:inc} by
constructing an atomic increment operation for a counter, like so:
\begin{code}
\begin{Verbatim}[xleftmargin=5mm,numbers=left]
def atomic_inc(ptr):
    atomic:
        ^ptr = ^ptr + 1;
    ;
;
count = 0;
atomic_inc(&count);
\end{Verbatim}
\end{code}
